Carbonate chemistry at Ny-Ålesund

### 2021-03-18 awipev-CO2.Rmd
choice of constants + plots work

#################### Regression function
# function regression plot with equation (MA or OLS) in title
ggreg <- function (xdata, ydata, type = "OLS", plot = TRUE, title = "title") { # x and y are the names of the variables
  fit_data <- tibble(xdata, ydata)
  colnames(fit_data) = c("xdata", "ydata")
  fit <- print(lmodel2::lmodel2(ydata ~ xdata))
if (type == "OLS") ty = 1 #one selects OLS only
if (type == "MA") ty = 2 #one selects MA only
#reg <- fit$regression.results[ty,] #one selects OLS only
intercept <- fit$regression.results$Intercept[ty]
slope <- fit$regression.results$Slope[ty]
if (plot == TRUE) {
  plt <- ggplot(data = fit_data, aes(x = xdata, y = ydata)) + 
  geom_point(na.rm = TRUE, size = 2, colour = "blue") +
  geom_abline(aes(intercept = intercept, slope = slope),
              colour = "blue")  + 
  labs(title = paste(title, "; type: ", type,
                     "\n R2 = ", round(fit$rsquare, digits = 2),
                     "; Intercept =", signif(intercept, 3),
                     "; Slope =", signif(slope, 3),
                     "; P =", signif(fit$P.param, 3)))
  return(plt)
  } else {
  return(fit)
}
}

@@

# Conversion of spec pH of in situ samples to in situ temp
pH_seafet <- d %>%
  dplyr::filter(!is.na(ph_s_sf) & !is.na(temp_insitu_ctd))
pH_seafet <- pH_seafet %>% 
  dplyr::mutate(
    ph_s_sf_temp_insi = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = pH_seafet$temp_insitu_ctd, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "x", kf = "x", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_sf_norm4 = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "x", kf = "x", ks="d", 
                         pHscale = "T", b="u74")
  )

@@

# Make mean for replicates + SD
pH_seafet_mean <- pH_seafet%>%
  dplyr::select(date, ph_s_sf_temp_insi)%>% 
  dplyr::group_by(date) %>% 
  dplyr::summarise(ph_s_sf_temp_insi_mean = mean(ph_s_sf_temp_insi, na.rm = TRUE),
                   ph_s_sf_temp_insi_sd= sd(ph_s_sf_temp_insi, na.rm = TRUE)
                   )

pH_durafet_mean <- pH_durafet%>%
  dplyr::select(date, ph_s_dur_temp_fb)%>% 
  dplyr::group_by(date) %>% 
  dplyr::summarize(ph_s_dur_temp_fb_mean = mean(ph_s_dur_temp_fb, na.rm = TRUE),
                   ph_s_dur_temp_fb_sd = sd(ph_s_dur_temp_fb, na.rm = TRUE))

#bind all spectro measurements, vhange name of dataframe only to do the rbind.
seafetph <- pH_seafet_mean
colnames(seafetph) <- c("date", "mean", "sd")
durafetph <- pH_durafet_mean
colnames(durafetph) <- c("date", "mean", "sd")
pH_seafet_durafet <- rbind(seafetph,durafetph)

#calculations for all pH spectro measurement for sd
sd_min <- min(pH_seafet_durafet$sd)
sd_max <- max(pH_seafet_durafet$sd)
sd_mean <- mean(pH_seafet_durafet$sd)

@@

# seaFET pH 
discrete2merge <- discrete %>%
  dplyr::select(closest_datetime, ph_s_sf_temp_insi_mean) %>%
  dplyr::rename(ph_s_sf_temp_insi = ph_s_sf_temp_insi_mean) %>%
  dplyr::filter(!is.na(closest_datetime)) %>%
  rename(datetime=closest_datetime)
d_all <- full_join(d_all, discrete2merge, by="datetime")
d_all <- d_all %>% 
  dplyr::mutate(ph_s_sf_temp_insi = ifelse(n_depth_ctd == 4, ph_s_sf_temp_insi, NA)) # eliminate ph spec data when the profiler was at several depth during the same hour
  #dplyr::mutate(ph_s_sf_temp_insi = ifelse(!is.na(temp_insitu_ctd_7m) |
  #                                           !is.na(temp_insitu_ctd_9m), 
   #                                        ph_s_sf_temp_insi, NA)) # only use ph spec data when profiler was below 6 m

@@

d_all <- d_all %>%
  dplyr::mutate(ph_s_sf_norm4 = ifelse(!is.na(pH_seafet$ph_s_sf),
                                              pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "x", kf = "x", ks="d", 
                         pHscale = "T", b="u74"), NA
  ))

@@

pH_sf$pHint_tot_sf_norm4 <- pHinsi(pH=pH_sf$pHint_tot_sf, ALK=mean(d_all$at, na.rm=TRUE), Tinsi=4, Tlab=pH_sf$temp_insitu_ctd_9m, Pinsi=11/10, S=pH_sf$sal_insitu_ctd_9m, Pt=0, Sit=0, k1k2 = "x", kf = "x", ks="d", pHscale = "T", b="u74")

pH_dur <- d_all %>%
  dplyr::select(datetime, sal_fb,temp_fb, ph_dur_corr)%>%
  dplyr::filter(!is.na(ph_dur_corr), !is.na(temp_fb), !is.na(sal_fb))
pH_dur$pH_durafet_norm4 <- pHinsi(pH=pH_dur$ph_dur_corr, ALK=mean(d_all$at, na.rm=TRUE), Tinsi=4, Tlab=pH_dur$temp_fb, Pinsi=11/10, S=pH_dur$sal_fb, Pt=0, Sit=0, k1k2 = "x", kf = "x", ks="d", pHscale = "T", b="u74")

@@

### **Calculated / Durafet pHs**
```{r pHs comparaison dur calc, echo=FALSE, message= FALSE, warning=FALSE}
# Normalize phcalc from AT/CT to 4°c
pH_calc <- discrete %>%
  dplyr::select(datetime, sal_fb, temp_fb, phcalc_fb, at)%>%
  dplyr::filter(!is.na(phcalc_fb), !is.na(temp_fb), !is.na(sal_fb))
pH_calc$pH_calc_fb_norm4 <- pHinsi(pH = pH_calc$phcalc_fb, ALK=pH_calc$at*1e-6, Tinsi=4, Tlab=pH_calc$temp_fb, Pinsi=11/10, S=pH_calc$sal_fb, Pt=0, Sit=0, k1k2 = "x", kf = "x", ks="d", pHscale = "T", b="u74")

@@

## Impact of choice of constants
### Lueker vs Sulpis using means of monthly means
```{r impact of constants 1, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis
# use mean values calculated for impacts of nutrients above

@@

```{r impact of constants 2, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis
tmp <- d_all %>% 
  drop_na(pco2_fb, at_calc, sal_fb, temp_fb)
carb_l <-
  carb(
    24,
    tmp$pco2_fb,
    tmp$at_calc * 1e-6,
    S = tmp$sal_fb,
    T = tmp$temp_fb,
    P = 0,
    Pt = 0,
    Sit = 0,
    k1k2 = "l",
    kf = "dg",
    ks = "d",
    pHscale = "T",
    b = "u74"
  )
carb_s <-
  carb(
    24,
    tmp$pco2_fb,
    tmp$at_calc * 1e-6,
    S = tmp$sal_fb,
    T = tmp$temp_fb,
    P = 0,
    Pt = 0,
    Sit = 0,
    k1k2 = "s20",
    kf = "dg",
    ks = "d",
    pHscale = "T",
    b = "u74"
  )
diff_ls <- as_tibble(carb_l - carb_s) %>% 
  dplyr::select(pH, pCO2, DIC, OmegaAragonite)
diff_ls_quant <- bind_rows(
  quantile(diff_ls$pH), quantile(diff_ls$pCO2), quantile(diff_ls$DIC * 1e6), 
  quantile(diff_ls$OmegaAragonite)) %>% 
  dplyr::mutate(Variable = c("pH", "pCO2", "DIC", "OmegaAragonite"), .before = 1)
pander(diff_ls_quant)

lm <- lm2(x = d_all$sal_insitu_ctd_9m, y = d_all$sal_fb)
p <-  ggplot(d_all,aes(x=sal_insitu_ctd_9m, y= sal_fb))+
  geom_point( aes(color=factor(monthd))) + 
  geom_abline(slope=1, intercept=0, linetype = "dashed")+
  geom_abline(aes(intercept = lm$fit$regression.results[2,2], slope = lm$fit$regression.results[2,3]),colour = "blue") +
  ylim(c(30.5,35.5)) +xlim(c(28,35.5)) +
  labs(title = paste("","\n Adj R2 = ", signif(lm$fit$rsquare, 3),
                     "; Intercept =", signif(lm$fit$regression.results[2,2], 3),
                     "; Slope =", signif(lm$fit$regression.results[2,3], 3),
                     "; P =", signif(lm$fit$P.param, 3),
                     "; RMSE = ", signif(lm$rmse, 3))) +
  theme(aspect.ratio = 1, plot.title = element_text(size = 7)) + #
  coord_fixed(ratio = 1 )+
  labs( x="Salinity CTD 9 m", y="Salinity FB ", color = "Months") + theme_bw()

@@

>>>>

#################### Regression function
# function regression plot with equation (MA or OLS) in title
ggreg <- function (xdata, ydata, type = "OLS", plot = TRUE, title = "title", subtitle = "subtitle", xlab = "x", ylab = "y") { # x and y are the names of the variables
  fit_data <- tibble(xdata, ydata)
  colnames(fit_data) = c("xdata", "ydata")
  fit <- print(lmodel2::lmodel2(ydata ~ xdata))
if (type == "OLS") ty = 1 #one selects OLS only
if (type == "MA") ty = 2 #one selects MA only
#reg <- fit$regression.results[ty,] #one selects OLS only
intercept <- fit$regression.results$Intercept[ty]
slope <- fit$regression.results$Slope[ty]
if (plot == TRUE) {
  plt <- ggplot(data = fit_data, aes(x = xdata, y = ydata)) + 
  geom_point(na.rm = TRUE, size = 2, colour = "blue") +
  geom_abline(aes(intercept = intercept, slope = slope),
              colour = "blue")  + 
  labs(title = paste(title, "; type: ", type,
                     "\n R2 = ", round(fit$rsquare, digits = 2),
                     "; Intercept =", signif(intercept, 3),
                     "; Slope =", signif(slope, 3),
                     "; P =", signif(fit$P.param, 3)),
       subtitle = subtitle,
       x = xlab,
       y = ylab)
  return(plt)
  } else {
  return(fit)
}
}

@@

# Conversion of spec pH of in situ samples to in situ temp and 4 °C
pH_seafet <- d %>%
  dplyr::filter(!is.na(ph_s_sf) & !is.na(temp_insitu_ctd))
pH_seafet <- pH_seafet %>% 
  dplyr::mutate(
    ph_s_sf_temp_insi = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = pH_seafet$temp_insitu_ctd, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "x", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_sf_norm4 = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "x", ks="d", 
                         pHscale = "T", b="u74")
  )

@@

# Make mean for replicates + SD
pH_seafet_mean <- pH_seafet%>%
  dplyr::select(date, ph_s_sf_temp_insi, ph_s_sf_norm4)%>% 
  dplyr::group_by(date) %>% 
  dplyr::summarise(ph_s_sf_temp_insi_mean = mean(ph_s_sf_temp_insi, na.rm = TRUE),
                   #ph_s_sf_temp_insi_sd = sd(ph_s_sf_temp_insi, na.rm = TRUE),
                   ph_s_sf_norm4_mean = mean(ph_s_sf_norm4, na.rm = TRUE)
                   )

pH_durafet_mean <- pH_durafet%>%
  dplyr::select(date, ph_s_dur_temp_fb)%>% 
  dplyr::group_by(date) %>% 
  dplyr::summarize(ph_s_dur_temp_fb_mean = mean(ph_s_dur_temp_fb, na.rm = TRUE)#,
                   #ph_s_dur_temp_fb_sd = sd(ph_s_dur_temp_fb, na.rm = TRUE)
                   )

# #bind all spectro measurements, vhange name of dataframe only to do the rbind.
# seafetph <- pH_seafet_mean
# colnames(seafetph) <- c("date", "ph_s_sf_insi", "ph_s_sf_norm4")
# durafetph <- pH_durafet_mean
# colnames(durafetph) <- c("date", "ph_s_dur_temp_fb")
# pH_seafet_durafet <- rbind(seafetph,durafetph)

# #calculations for all pH spectro measurement for sd
# sd_min <- min(pH_seafet_durafet$sd)
# sd_max <- max(pH_seafet_durafet$sd)
# sd_mean <- mean(pH_seafet_durafet$sd)

@@

# seaFET pH 
discrete2merge <- discrete %>%
  dplyr::select(closest_datetime, ph_s_sf_temp_insi_mean, ph_s_sf_norm4_mean) %>%
  dplyr::rename(ph_s_sf_temp_insi = ph_s_sf_temp_insi_mean,
                ph_s_sf_norm4 = ph_s_sf_norm4_mean) %>%
  dplyr::filter(!is.na(closest_datetime)) %>%
  rename(datetime=closest_datetime)
d_all <- full_join(d_all, discrete2merge, by="datetime")
d_all <- d_all %>% 
  dplyr::mutate(ph_s_sf_temp_insi = ifelse(n_depth_ctd == 4, ph_s_sf_temp_insi, NA)) # eliminate ph spec data when the profiler was at several depth during the same hour
  #dplyr::mutate(ph_s_sf_temp_insi = ifelse(!is.na(temp_insitu_ctd_7m) |
  #                                           !is.na(temp_insitu_ctd_9m), 
   #                                        ph_s_sf_temp_insi, NA)) # only use ph spec data when profiler was below 6 m

@@

# d_all <- d_all %>%
#   dplyr::mutate(ph_s_sf_norm4 = ifelse(!is.na(pH_seafet$ph_s_sf),
#                                               pHinsi(pH = pH_seafet$ph_s_sf, 
#                          ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = 4, 
#                          Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, 
#                          S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "x", ks="d", 
#                          pHscale = "T", b="u74"), NA
#   ))

@@

pH_sf$pHint_tot_sf_norm4 <- pHinsi(pH=pH_sf$pHint_tot_sf, ALK=mean(d_all$at, na.rm=TRUE), Tinsi=4, Tlab=pH_sf$temp_insitu_ctd_9m, Pinsi=11/10, S=pH_sf$sal_insitu_ctd_9m, Pt=0, Sit=0, k1k2 = "l", kf = "x", ks="d", pHscale = "T", b="u74")

pH_dur <- d_all %>%
  dplyr::select(datetime, sal_fb,temp_fb, ph_dur_corr)%>%
  dplyr::filter(!is.na(ph_dur_corr), !is.na(temp_fb), !is.na(sal_fb))
pH_dur$pH_durafet_norm4 <- pHinsi(pH=pH_dur$ph_dur_corr, ALK=mean(d_all$at, na.rm=TRUE), Tinsi=4, Tlab=pH_dur$temp_fb, Pinsi=11/10, S=pH_dur$sal_fb, Pt=0, Sit=0, k1k2 = "l", kf = "x", ks="d", pHscale = "T", b="u74")

@@

### **Calculated / Durafet pHs**
```{r pHs comparaison dur calc, echo=FALSE, message= FALSE, warning=FALSE}
# Normalize phcalc from AT/CT to 4°c
pH_calc <- discrete %>%
  dplyr::select(datetime, sal_fb, temp_fb, phcalc_fb, at)%>%
  dplyr::filter(!is.na(phcalc_fb), !is.na(temp_fb), !is.na(sal_fb))
pH_calc$pH_calc_fb_norm4 <- pHinsi(pH = pH_calc$phcalc_fb, ALK=pH_calc$at*1e-6, Tinsi=4, Tlab=pH_calc$temp_fb, Pinsi=11/10, S=pH_calc$sal_fb, Pt=0, Sit=0, k1k2 = "l", kf = "x", ks="d", pHscale = "T", b="u74")

@@

## Impact of choice of constants  {.tabset .tabset-pills}
### Lueker vs Sulpis using means of monthly means
```{r impact of constants 1, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis
# use mean values calculated for impacts of nutrients above

@@

```{r impact of constants 2, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis using pco2_fb and at_calc
tmp <- d_all %>% 
  tidyr::drop_na(pco2_fb, at_calc, sal_fb, temp_fb) %>% 
  dplyr::select(ph_s_sf_norm4, pco2_fb, at_calc, sal_fb, temp_fb)

carb_l <-
  carb(
    24,
    tmp$pco2_fb,
    tmp$at_calc * 1e-6,
    S = tmp$sal_fb,
    T = tmp$temp_fb,
    P = 0,
    Pt = 0,
    Sit = 0,
    k1k2 = "l",
    kf = "dg",
    ks = "d",
    pHscale = "T",
    b = "u74"
  )
# normalize derived pH at 4°C
carb_l$pH <- pHinsi(pH = carb_l$pH, 
                         ALK = carb_l$ALK, Tinsi = 4, 
                         Tlab = carb_l$T, Pinsi = 0, 
                         S = carb_l$S, Pt = 0, Sit = 0, k1k2 = "l", kf = "x", ks="d", 
                         pHscale = "T", b="u74")
carb_s <-
  carb(
    24,
    tmp$pco2_fb,
    tmp$at_calc * 1e-6,
    S = tmp$sal_fb,
    T = tmp$temp_fb,
    P = 0,
    Pt = 0,
    Sit = 0,
    k1k2 = "s20",
    kf = "dg",
    ks = "d",
    pHscale = "T",
    b = "u74"
  )
# normalize derived pH at 4°C
carb_s$pH <- pHinsi(pH = carb_s$pH, 
                         ALK = carb_s$ALK, Tinsi = 4, 
                         Tlab = carb_s$T, Pinsi = 0, 
                         S = carb_s$S, Pt = 0, Sit = 0, k1k2 = "l", kf = "x", ks="d", 
                         pHscale = "T", b="u74")

# do difference
diff_ls <- as_tibble(carb_l - carb_s) %>% 
  dplyr::rename(delta_ph = pH, delta_ct = DIC, delta_OmegaAragonite = OmegaAragonite) %>%
  dplyr::select(delta_ph, delta_ct, delta_OmegaAragonite)

# calculate quantiles
diff_ls_quant <- bind_rows(
  quantile(diff_ls$delta_ph), quantile(diff_ls$delta_ct * 1e6),
  quantile(diff_ls$delta_OmegaAragonite)) %>% 
  dplyr::mutate(Variable = c("pH", "DIC", "OmegaAragonite"), .before = 1)
#pander(diff_ls_quant)

# prepare tibble to compare calculated pH and spec pH
tmp2 <- tibble(ph_s_sf_norm4 = tmp$ph_s_sf_norm4, pH_l = carb_l$pH, pH_s = carb_s$pH) %>% 
  tidyr::drop_na(ph_s_sf_norm4)

p1 <- ggreg(xdata = tmp2$ph_s_sf_norm4, ydata = tmp2$pH_l, type = "OLS", plot = TRUE, 
            title = "both pH normalized at 4°C", subtitle = "",
            xlab = "Spectrophotometric pH measured @ 4°C",
            ylab = "pH @ 4°C calculated (pCO2+AT) with Lueker")
p2 <- ggreg(xdata = tmp2$ph_s_sf_norm4, ydata = tmp2$pH_s, type = "OLS", plot = TRUE, title = "both pH normalized at 4°C")

lm <- lm2(x = tmp2$ph_s_sf_norm4, y = tmp2$pH_l)
p <-  ggplot(d_all,aes(x=sal_insitu_ctd_9m, y= sal_fb))+
  geom_point( aes(color=factor(monthd))) + 
  geom_abline(slope=1, intercept=0, linetype = "dashed")+
  geom_abline(aes(intercept = lm$fit$regression.results[2,2], slope = lm$fit$regression.results[2,3]),colour = "blue") +
  ylim(c(30.5,35.5)) +xlim(c(28,35.5)) +
  labs(title = paste("","\n Adj R2 = ", signif(lm$fit$rsquare, 3),
                     "; Intercept =", signif(lm$fit$regression.results[2,2], 3),
                     "; Slope =", signif(lm$fit$regression.results[2,3], 3),
                     "; P =", signif(lm$fit$P.param, 3),
                     "; RMSE = ", signif(lm$rmse, 3))) +
  theme(aspect.ratio = 1, plot.title = element_text(size = 7)) + #
  coord_fixed(ratio = 1 )+
  labs( x="Salinity CTD 9 m", y="Salinity FB ", color = "Months") + theme_bw()

@@

`r knitr::kable(diff_ls_quant, digits = c(0, rep(3, 5)), caption = "Lueker minus Sulpis", format = "simple")`
